---
title: "Course Project"
author: "ED"
date: "7/2/2020"
output: github_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("WorkSpace.RData")
library(dplyr)
library(knitr)
```
This code book describes the source of the data, variables and transformations done to create a tidy data set. The project represents the final assignment for the Getting and Cleaning Data course by John Hopkins University in Coursera.  
  
### Data
The data used for this project was obtained from the following site:
<https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>

The data comes from the experiment of Human Activity Recognition Ver 1.0 published in 2012 by Smartlab, which represents data collected and derived from the accelerometer and gyroscope of smartophone Samsung Galaxy S II. The 'UCI HAR Dataset/README.txt' file within 'Dataset.zip' provides the full description of the raw data set and included files structure.  

### Variables
The raw data includes 561 of combined time and frequency domain variables from the accelerometer and gyroscope 3-axial raw signals. The full description of such variables are explained in 'UCI HAR Dataset/features_info.txt'.

Two additional numeric variables are provided that capture the activity label and the subject who performed the activity for each observation in the data set.

### Transformation 
The code for the steps describes here are recorded in script `run_analysis.R`

1. The training (7352 obs.) and the test (2947 obs.) sets are merged to create one data set named 'ds' of dimension: 
```{r echo=FALSE}
dim(ds)
```
  
  
2. 'ds' variables are renamed using 'UCI HAR Dataset/features.txt'; which allow to select the measurements on the mean and standard deviation. The new subset is saved as 'ds1' of dimension:
```{r echo=FALSE}
dim(ds1)
```
  
  
3. The training (7352 obs.) and test (2947 obs.) numeric labels are joined together maintaining the same index order as 'ds' in step 1. Subsequently, the activity legend key in 'UCI HAR Dataset/activity_labels.txt' is used to to extract the descriptive activity names which are added to 'ds1' under variable name 'activity'. The resultant unique count per activity names in 'ds1' are:
```{r echo=FALSE, message=FALSE}
x <- ds1 %>% group_by(activity) %>% summarise(count = n()) 
kable(x)
```
  
  
4. The variable names of data set 'ds1' are cleaned up to a more descriptive format, removing all special characters, replacing ^t and ^f with "time" and "frequency"; and finally converting all upper to lower case.  
        i. `names(ds1[1]) "tBodyAcc-mean()-X"` *original*    
        ii. `names(ds1[1]) "timeBodyAccmeanX"`  
        iii. `names(ds1[1])"timebodyaccmeanx"` *final*    

        
5. The numeric subject data for training (7352 obs.) and test (2947 obs.) are joined together maintaining the same index order as 'ds' in step 1; and then merged with 'ds1'. From data set 'ds1' a second, independent tidy data set 'ds2'is created which contains the average of each variable for each activity and each subject, which dimension is:

```{r echo=FALSE}
dim(ds2)
```

'ds2' with 188 variables can be difficult to visualize in the console: 
```{r}
head(ds2[,1:10])
```
  
  
hence a vertical version of 'ds2' is also offered 'ds2ver':
```{r echo=FALSE}
head(ds2ver)
```

